---
title: "US Helicoverpa BYM"
output:
  html_document:
    toc: true
    toc_depth: 2
    toc_float: true
    number_sections: true
    theme: sandstone
    highlight: tango
    code_folding: show
editor_options: 
  chunk_output_type: console
---

```{r}
setwd("/Users/ddlawton/Dropbox (ASU)/Research/Postdoc/Helicoverpa_climatic_variability")
```
```{r}

#Main modeling package
suppressMessages(library(INLA))
#install.packages("INLA",repos=c(getOption("repos"),INLA="https://inla.r-inla-download.org/R/stable"), dep=TRUE)


#supporting packages
suppressMessages(library(reshape2))
suppressMessages(library(raster))
suppressMessages(library(corrplot))
suppressMessages(library(rasterVis))
suppressMessages(library(rgeos))
suppressMessages(library(rgl))
suppressMessages(library(maptools))
suppressMessages(library(mapproj))
suppressMessages(library(spdep))
suppressMessages(library(ggplot2))
suppressMessages(library(ggthemes))
suppressMessages(library(GISTools))
suppressMessages(library(lattice))
suppressMessages(library(gridExtra))
suppressMessages(library(spatialEco))
suppressMessages(library(dplyr))
suppressMessages(library(sp))
suppressMessages(library(pals))
suppressMessages(library(viridis)) #only new package from mesh code
suppressMessages(library(devtools))
#install_github('timcdlucas/INLAutils')
suppressMessages(library(INLAutils))
suppressMessages(library(splines))

```

# Observations  
```{r}
Heli.df = read.csv("data/processed/US_seasonal_climate.csv", 
                  header = TRUE, sep=",")

dim(Heli.df)
str(Heli.df)

range(Heli.df$Year)
unique(Heli.df$Species)

range(Heli.df$Latitude)
range(Heli.df$Longitude)

range(Heli.df$Num_traps) #sampling effort

range(Heli.df$mean_count)
length(which(is.na(Heli.df$se_count)))
length(which(is.na(Heli.df))) #all missing/NA values are in se_count
```


## Observations to Spatial Points
```{r}
LL84 = "+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0"

Heli.pnts = SpatialPointsDataFrame(Heli.df[,c("Longitude","Latitude")], Heli.df)
proj4string(Heli.pnts) = LL84

plot(Heli.pnts, col="red", cex=0.1)
```


# Option 2 (BYM)  
Besag_York_Mollie model using areal units. Example for US. 

## Spatial Extents
Load US State boundaries
```{r}
LL84 = "+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0"

nProj = "+proj=merc +a=6378137 +b=6378137 +lat_ts=0 +lon_0=0
         +x_0=0 +y_0=0 +k=1 +units=km +nadgrids=@null +no_defs"

#States
States = map("state", 
            fill = TRUE,
            plot = FALSE)

IDs = sapply(strsplit(States$names, ":"),
             function(x) x[1])

States = map2SpatialPolygons(States, IDs = IDs,
                              proj4string = CRS(LL84))
 
pid = sapply(slot(States, "polygons"), 
             function(x) slot(x, "ID"))

p.df = data.frame(ID=1:length(States), 
                  row.names = pid)

States = SpatialPolygonsDataFrame(States, p.df)

States.merge = gUnaryUnion(States) #merge states to create US boundaries.  Ignore polygon toplogy warnings, this is just for mapping

plot(States, col = "gray90") #Check States
plot(States.merge, border = "red", lwd = 2, add=T) #merged

```


## Construct Hexagons    
Construct a hexagon frame around study area to define areal regions.
```{r warning=FALSE}
States.buff = gBuffer(States.merge, #Create a buffer so hexagon extend just outside of US
                     width = 0.25, #quarter degree
                     byid = FALSE) 


#First need to make a regular point grid; these will be hexagon centroids
size = 1 #controls spacing/resolution. Smaller value = more points
set.seed(1976)
hex_points = spsample(States.buff,
                      type = "hexagonal", cellsize = size)

#Next, use points to create hexagons
hex_grid = HexPoints2SpatialPolygons(hex_points, dx = size)

#Add a dataframe to grid 
pid = sapply(slot(hex_grid, "polygons"), 
             function(x) slot(x, "ID"))

p.df = data.frame(ID=1:length(hex_grid), 
                  row.names = pid)

hex_grid = SpatialPolygonsDataFrame(hex_grid, p.df)

dim(hex_grid)[1] # number of hexagons.  For comparison, there are about 3000 counties in US

#Add ID number, calling cells "Regions" 
hex_grid$Region = 1:nrow(hex_grid@data)

#Project cooridnates
hex_grid = spTransform(hex_grid, nProj)

#Calculate area
hex_grid$Area = gArea(hex_grid, byid = TRUE)

#range, km2
range(hex_grid$Area)
```


## Select and Project US Observations
```{r}
Heli.pnts = spTransform(Heli.pnts, proj4string(hex_grid)) #project to same as hexagons
dim(Heli.pnts)[1] #Total points

#Filter observations to US only
Heli.pnts$Region = over(Heli.pnts, hex_grid)[,"Region"] #Extract corresponding hexagon region
Heli.pnts = subset(Heli.pnts, is.na(Region) == FALSE) # filter to only moths in US

range(Heli.pnts$Region) #Region IDs

dim(Heli.pnts)[1] #US only points

#Quick Check
plot(hex_grid) #Verify US only
plot(Heli.pnts, add=T, col="red", cex=0.5)
```


## Region Counts   
Count number of observations and traps in each region by year.
```{r}
Region.counts = as.data.frame(
                      Heli.pnts@data %>%
                        group_by(Year, Region) %>%
                        summarise(Moths = sum(mean_count),
                                  num_traps = length(unique(Trap)),
                                  Air_temp = first(Air_temp_mean),
                                  Precip = first(Precip_mean)))



head(Region.counts)
summary(Region.counts)
```

## Map Results
Mapping 2010 counts to check results.  Looks like very observations for 2021
```{r  fig.width=8, fig.height=15}
plot.2010 = Region.counts %>% filter(Year == 2010) #filter to 2021?

hex_grid.plot = hex_grid #mae copy for plotting

hex_grid.plot@data = left_join(hex_grid.plot@data, plot.2010, by = "Region") #match by region ID

#Not all hexagons had moths, these location have NA, this can be ignored for plotting now, but may need changed later (set to zero?)
head(hex_grid.plot)


#Format for ggplot
hex_grid.plot@data$id = rownames(hex_grid.plot@data)
hex_df = fortify(hex_grid.plot, region = "id")
hex_df = left_join(hex_df, hex_grid.plot@data, by = "id")


ggplot(hex_df, 
        aes(long,lat, group=group, fill = Moths)) + 
        scale_fill_viridis(name="Moth Counts 2010", 
                           discrete=F, option = "inferno",
                           direction = -1, na.value = "white") +
       geom_polygon(col="black", size = 0.1) + 
        xlab(" ") +
        ylab(" ") +
        coord_equal() + 
        theme(panel.grid.minor = element_blank(),
              panel.grid.major = element_blank(),
              panel.background = element_blank(),
              plot.background = element_blank(),
              panel.border = element_blank(),
              #legend.direction = "horizontal",
              legend.position="bottom",
              strip.text = element_text(size=16, face="bold"),
              strip.background = element_blank(),
              legend.key.size = unit(1,"line"),
              legend.key.width = unit(3,"line"),
              legend.text = element_text(size=10, face="bold"),
              legend.title = element_text(size=14, face="bold"),
              axis.title.x = element_blank(),
              axis.title.y =element_blank(),
              axis.text.x = element_blank(),
              axis.text.y =element_blank(),
              axis.ticks.x = element_blank(),
              axis.ticks.y =element_blank(),
              plot.title = element_text(size=22, face="bold"))


ggplot(hex_df, 
        aes(long,lat, group=group, fill = num_traps)) + 
        scale_fill_viridis(name="Trap Counts 2010", 
                           discrete=F, option = "viridis",
                           direction = -1, na.value = "white") +
       geom_polygon(col="black", size = 0.1) + 
        xlab(" ") +
        ylab(" ") +
        coord_equal() + 
        theme(panel.grid.minor = element_blank(),
              panel.grid.major = element_blank(),
              panel.background = element_blank(),
              plot.background = element_blank(),
              panel.border = element_blank(),
              #legend.direction = "horizontal",
              legend.position="bottom",
              strip.text = element_text(size=16, face="bold"),
              strip.background = element_blank(),
              legend.key.size = unit(1,"line"),
              legend.key.width = unit(3,"line"),
              legend.text = element_text(size=10, face="bold"),
              legend.title = element_text(size=14, face="bold"),
              axis.title.x = element_blank(),
              axis.title.y =element_blank(),
              axis.text.x = element_blank(),
              axis.text.y =element_blank(),
              axis.ticks.x = element_blank(),
              axis.ticks.y =element_blank(),
              plot.title = element_text(size=22, face="bold"))
```



## Expand Dataset
For each year in the analysis, we need to identify which hexagons do NOT have moths.  Here, we loop through each year so that all regions are accounted for.
```{r}
Year.levels = unique(Region.counts$Year) 
range(Year.levels) #years in analysis

for(i in 1:length(Year.levels)){ #For each year i...
  
     tmp.obs = Region.counts %>% filter(Year == Year.levels[i]) # filter observations for that year
     
     tmp.hex = hex_grid@data #Make a year specific copy of the hexagon data
     tmp.hex$Year = Year.levels[i] #label year in this set of a hexagons
     
     tmp.df = left_join(tmp.hex, tmp.obs, by= c("Year","Region")) #match observations to hexagons
     
     if(i == 1){Model.hex = tmp.df #stack each year on top
     } else{Model.hex = rbind(Model.hex, tmp.df)}
 }
       
head(Model.hex) 
Model.hex$Moths[is.na(Model.hex$Moths)] = 0 #set cells without moths to 0
Model.hex$Traps[is.na(Model.hex$num_traps)] = 0 #same for traps

dim(Model.hex)
apply(Model.hex,2,range) #range of values
```


## Spatial Neighborhood Graph 
To estimate spatial relationships, need to identify regions sharing boundaries.  
```{r}
nb = poly2nb(hex_grid,  #identify neighbors from spatial polygons
             queen = TRUE) #consider all directions around cells

nb2INLA("J", nb) #format for modeling
J = INLA::inla.read.graph("J") #coverting to a matrix


summary(nb) # the header info is what is important: 995 cells, 5606 neighbors, an average of 5.6 neighbors per hexagon
```
     
## View Neighborhood graph
Graph identifies bordering cells.  Hexagons in background, centroids as points, and lines show cells that are neighbors.  

```{r fig.width=11, fig.height=11}
plot(hex_grid, col='gray95', border='gray50')
xy = coordinates(hex_grid)
plot(nb, xy, col='red', lwd=0.15, add=TRUE)
```



## Model Setup

Need to create indices for temporal random effects.

### Year Index
```{r}
Model.hex$Year.int = as.integer(as.factor(as.character(Model.hex$Year)))

range(Model.hex$Year) #actual years
range(Model.hex$Year.int) #converted to index, year one through year 32
```

### Space-Time Interaction Index  
The relationship of space to time (region-years).  Basically, a variable for each possible Region and Year combination.
```{r}
Model.hex$Region.Yr = paste("ID", Model.hex$Region, "Y", Model.hex$Year, sep="")
head(Model.hex$Region.Yr) #like factor levels, where names include the Year and Region for all combinations

Model.hex$ID.Region.Yr = as.integer(as.factor(Model.hex$Region.Yr)) #convert the above to an index

range(Model.hex$ID.Region.Yr) #number of Region*Year combinations
```

### Vizualizing the climatic variables. 
Just temperature and precipitation for now.
```{r}
summary(Model.hex)
range(Model.hex$Air_temp,na.rm=TRUE)
range(Model.hex$Precip,na.rm=TRUE)

ggplot(Model.hex,aes(x=Air_temp)) + geom_histogram() + ggpubr::theme_pubr()

ggplot(Model.hex,aes(x=Precip)) + geom_histogram() + ggpubr::theme_pubr()

ggplot(Model.hex,aes(x=Moths)) + geom_histogram() + ggpubr::theme_pubr()


```

### Organize Data as List Object  
This is same as with the mesh approach, just organizing everything as a list() instead of a dataframe.  
```{r}
Model.hex$range.grp <- inla.group(Model.hex$Air_temp, n = 20, method = "quantile")


My.lst =  list(list(intercept1 = rep(1, dim(Model.hex)[1])), #Repeat value of 1 for each data point
          list(Air_temp = Model.hex[,"Air_temp"], # Seasonal average temperature
               Precip = Model.hex[,"Precip"], #Fake "fixed" effect
               Effort = Model.hex[,"num_traps"], # number of traps in hexagon, "Effort"
               Year1 = Model.hex[,"Year.int"], #Copy of Year index
               Year2 = Model.hex[,"Year.int"], #Extra copy of Year index, depending on how model is setup, sometimes need extra copied
               Year3 = Model.hex[,"Year.int"], #Extra copy of Year index
               Region1 = Model.hex[,"Region"], #Region/Hexagon identifier
               Region2 = Model.hex[,"Region"], #Extra copy of Region/Hexagon identifier
               Region3 = Model.hex[,"Region"], #Extra copy of Region/Hexagon identifier
               range.grp = Model.hex[,"range.grp"],
               ID.Region.Yr1  = Model.hex[,"ID.Region.Yr"])) #Region*Year combinations, "interaction"


range(Model.hex$Moths) #Very high values...  Need to think about how best to approach this issue, for now just creating a cutoff

Q.cut = quantile(Model.hex$Moths, probs = c(0.99)) #Just for this code example, id the 99th count of moths 
Q.cut #Cut point

length(which(Model.hex$Moths > Q.cut)) #number of regions with more than the cut off point

Model.hex$Moths.cut = ifelse(Model.hex$Moths >= Q.cut, Q.cut, Model.hex$Moths) #making Q.cut the maximum count, again just for code example

Model.hex$Moths.cut = round(Model.hex$Moths.cut, 0) #rounding to use poisson for demo
Model.hex$Moths2 = round(Model.hex$Moths, 0)

ggplot(Model.hex,aes(x=Moths,y=Moths.cut)) + geom_point()


summary(Model.hex$Moths) #Updated range

Stack.1 = inla.stack(data = list(Y = Model.hex$Moths2),  #Response variable, Moth Counts w/ added cut point
                                 A = list(1,1), # nothing here, just a placeholder for a possible matrix
                           effects = My.lst ,   #all predictors and space time indices above  
                               tag = "area.0")  #arbitrary label/name


Stack.2 = inla.stack(data = list(Y = Model.hex$Moths2),  #Response variable, Moth Counts w/ added cut point
                                 A = list(1,1), # nothing here, just a placeholder for a possible matrix
                           effects = My.lst ,   #all predictors and space time indices above  
                               tag = "area.0")  #arbitrary label/name


```


## Run Model
```{r}
Frm.0 = Y ~ -1 + intercept1 +      #Y is response variable, -1 tells software we're using a custom intercept1
                       f(Region1,  #Index of Region IDs
                         model="bym2", #specifying we want a BYM (Besag Mollie York) model
                         graph=J,      #J is the matrix that identfies neighboring hexagons
                         constr=TRUE,  #center/scale the random spatial effect to zero
                         hyper=list(phi = list(prior = "pc",     #priors that inform the spatial effect, we can dive into these later.
                                               param = c(0.5, 2/3), 
                                               initial = 3), 
                                    prec = list(prior = "pc.prec", 
                                                param = c(1, 0.01), 
                                                initial = 1.5))) +
                        f(Year1,     #Year index
                          model="rw1", # model as a random walk, smooth trend
                          constr=TRUE) + #center on zero
                        f(ID.Region.Yr1,  #interaction effect
                          model="iid",  #each combination may vary
                          constr=TRUE) +
                       Effort 


inla.stack.data(Stack.2) <- inla.group(lidar$range, n = 20, method = "quantile")


Frm.1 = Y ~ -1 + intercept1 +      #Y is response variable, -1 tells software we're using a custom intercept1
                       f(Region1,  #Index of Region IDs
                         model="bym2", #specifying we want a BYM (Besag Mollie York) model
                         graph=J,      #J is the matrix that identfies neighboring hexagons
                         constr=TRUE,  #center/scale the random spatial effect to zero
                         hyper=list(phi = list(prior = "pc",     #priors that inform the spatial effect, we can dive into these later.
                                               param = c(0.5, 2/3), 
                                               initial = 3), 
                                    prec = list(prior = "pc.prec", 
                                                param = c(1, 0.01), 
                                                initial = 1.5))) +
                        f(Year1,     #Year index
                          model="rw2", # model as a random walk, smooth trend
                          constr=TRUE) + #center on zero
                        f(ID.Region.Yr1,  #interaction effect
                          model="iid",  #each combination may vary
                          constr=TRUE) +
                        f(range.grp,     #Year index
                          model="rw2", # model as a random walk, smooth trend
                          constr=TRUE)

Frm.2 = Y ~ -1 + intercept1 +      #Y is response variable, -1 tells software we're using a custom intercept1
                       f(Region1,  #Index of Region IDs
                         model="bym2", #specifying we want a BYM (Besag Mollie York) model
                         graph=J,      #J is the matrix that identfies neighboring hexagons
                         constr=TRUE,  #center/scale the random spatial effect to zero
                         hyper=list(phi = list(prior = "pc",     #priors that inform the spatial effect, we can dive into these later.
                                               param = c(0.5, 2/3), 
                                               initial = 3), 
                                    prec = list(prior = "pc.prec", 
                                                param = c(1, 0.01), 
                                                initial = 1.5))) +
                        f(Year1,     #Year index
                          model="rw2", # model as a random walk, smooth trend
                          constr=TRUE) + #center on zero
                        f(ID.Region.Yr1,  #interaction effect
                          model="iid",  #each combination may vary
                          constr=TRUE) +
                        ns(Air_temp,     #Year index
                          df=20)

Frm.3 = Y ~ -1 + intercept1 +      #Y is response variable, -1 tells software we're using a custom intercept1
                       f(Region1,  #Index of Region IDs
                         model="bym2", #specifying we want a BYM (Besag Mollie York) model
                         graph=J,      #J is the matrix that identfies neighboring hexagons
                         constr=TRUE,  #center/scale the random spatial effect to zero
                         hyper=list(phi = list(prior = "pc",     #priors that inform the spatial effect, we can dive into these later.
                                               param = c(0.5, 2/3), 
                                               initial = 3), 
                                    prec = list(prior = "pc.prec", 
                                                param = c(1, 0.01), 
                                                initial = 1.5))) +
                        f(Year1,     #Year index
                          model="rw2", # model as a random walk, smooth trend
                          constr=TRUE) + #center on zero
                        f(ID.Region.Yr1,  #interaction effect
                          model="iid",  #each combination may vary
                          constr=TRUE) +
                        Air_temp


knots <- seq(10, 30, by = 1)

Frm.4 = Y ~ -1 + intercept1 +      #Y is response variable, -1 tells software we're using a custom intercept1
                       f(Region1,  #Index of Region IDs
                         model="bym2", #specifying we want a BYM (Besag Mollie York) model
                         graph=J,      #J is the matrix that identfies neighboring hexagons
                         constr=TRUE,  #center/scale the random spatial effect to zero
                         hyper=list(phi = list(prior = "pc",     #priors that inform the spatial effect, we can dive into these later.
                                               param = c(0.5, 2/3), 
                                               initial = 3), 
                                    prec = list(prior = "pc.prec", 
                                                param = c(1, 0.01), 
                                                initial = 1.5))) +
                        f(Year1,     #Year index
                          model="rw2", # model as a random walk, smooth trend
                          constr=TRUE) + #center on zero
                        f(ID.Region.Yr1,  #interaction effect
                          model="iid",  #each combination may vary
                          constr=TRUE) +
                        bs(Air_temp, knots = knots)

                 

#theta1 = Model.0$internal.summary.hyperpar$mean
#theta1 = c(-3.8205006, 3.0879798, 0.6057112, -2.4115143) #mean parameters from first run


```


```{r}
Model.0 = inla(Frm.0, #formula above
      				 data = inla.stack.data(Stack.2), #data to use
      				 family = "poisson", #family
      				 verbose = TRUE, #print results to screen while running
      				 control.predictor = list( 
      											 A = inla.stack.A(Stack.1), #data again
      									 compute = TRUE, #estimate Fitted values using link function
      										  link = 1),  #use default link function, log-link for poisson 
      				 #control.mode = list(restart = TRUE, theta = theta1), #parameter estimates to speed things up 
      				 control.inla = list(strategy="gaussian", #speeds up run
      											   int.strategy = "eb"),
      				 control.compute=list(dic = TRUE, cpo = TRUE, waic = TRUE)) #calculate model comparison metrics
```

```{r}
Model.1 = inla(Frm.1, #formula above
      				 data = inla.stack.data(Stack.2), #data to use
      				 family = "poisson", #family
      				 verbose = TRUE, #print results to screen while running
      				 control.predictor = list( 
      											 A = inla.stack.A(Stack.1), #data again
      									 compute = TRUE, #estimate Fitted values using link function
      										  link = 1),  #use default link function, log-link for poisson 
      				 #control.mode = list(restart = TRUE, theta = theta1), #parameter estimates to speed things up 
      				 control.inla = list(strategy="gaussian", #speeds up run
      											   int.strategy = "eb"),
      				 control.compute=list(dic = TRUE, cpo = TRUE, waic = TRUE)) #calculate model comparison metrics
```


```{r}
Model.2 = inla(Frm.2, #formula above
      				 data = inla.stack.data(Stack.2), #data to use
      				 family = "poisson", #family
      				 verbose = TRUE, #print results to screen while running
      				 control.predictor = list( 
      											 A = inla.stack.A(Stack.1), #data again
      									 compute = TRUE, #estimate Fitted values using link function
      										  link = 1),  #use default link function, log-link for poisson 
      				 #control.mode = list(restart = TRUE, theta = theta1), #parameter estimates to speed things up 
      				 control.inla = list(strategy="gaussian", #speeds up run
      											   int.strategy = "eb"),
      				 control.compute=list(dic = TRUE, cpo = TRUE, waic = TRUE)) #calculate model comparison metrics
```


```{r}
Model.3 = inla(Frm.3, #formula above
      				 data = inla.stack.data(Stack.2), #data to use
      				 family = "poisson", #family
      				 verbose = TRUE, #print results to screen while running
      				 control.predictor = list( 
      											 A = inla.stack.A(Stack.1), #data again
      									 compute = TRUE, #estimate Fitted values using link function
      										  link = 1),  #use default link function, log-link for poisson 
      				 #control.mode = list(restart = TRUE, theta = theta1), #parameter estimates to speed things up 
      				 control.inla = list(strategy="gaussian", #speeds up run
      											   int.strategy = "eb"),
      				 control.compute=list(dic = TRUE, cpo = TRUE, waic = TRUE)) #calculate model comparison metrics
```

```{r}
Model.4 = inla(Frm.4, #formula above
      				 data = inla.stack.data(Stack.2), #data to use
      				 family = "poisson", #family
      				 verbose = TRUE, #print results to screen while running
      				 control.predictor = list( 
      											 A = inla.stack.A(Stack.1), #data again
      									 compute = TRUE, #estimate Fitted values using link function
      										  link = 1),  #use default link function, log-link for poisson 
      				 #control.mode = list(restart = TRUE, theta = theta1), #parameter estimates to speed things up 
      				 control.inla = list(strategy="gaussian", #speeds up run
      											   int.strategy = "eb"),
      				 control.compute=list(dic = TRUE, cpo = TRUE, waic = TRUE)) #calculate model comparison metrics
```


```{r}
Model.0$dic$dic
Model.1$dic$dic
Model.2$dic$dic
Model.3$dic$dic
Model.4$dic$dic


Model.0$waic$waic
Model.1$waic$waic
Model.2$waic$waic
Model.3$waic$waic
Model.4$waic$waic

```


Results summary
```{r}
sum_mod0 <- summary(Model.0) #mean result values, specific values are plotted below 
sum_mod01 <- summary(Model.1) #mean result values, specific values are plotted below 

sum_mod0$
sum_mod01$dic

library(INL)
INLAdispersion_check(Model.1, nsim = 1000)

```

## Check Fitted Values
```{r  fig.width=11, fig.height=8}
#Pull fitted values
idat = inla.stack.index(Stack.1, "area.0")$data    #Create an index corresponding to input data, "area.0" is the arbitrary name in the input data (Stack.1)

idat2 = inla.stack.index(Stack.2, "area.0")$data    #Create an index corresponding to input data, "area.0" is the arbitrary name in the input 
Model.hex$Fitted = Model.0$summary.fitted.values$mean[idat]  #Get mean fitted values from model
Model.hex$Fitted2 = Model.1$summary.fitted.values$mean[idat2]  #Get mean fitted values from model


plot.fitted = Model.hex %>% filter(Moths > 0) # remove all the extra zeros to compare fitted against observed moths
cor(plot.fitted$Moths2, plot.fitted$Fitted) #rounder version with maximum cutoff
cor(plot.fitted$Moths, plot.fitted$Fitted) #original values before rounding, cutoff, etc..


ggplot(plot.fitted, aes(Moths2, Fitted)) +
         geom_jitter(shape = 1, size = 3, col = "gray70") +
         geom_abline(slope = 1, intercept = 0, alpha = 0.5, size = 1, col="red") +
         xlab("Observed Moth Counts") +
            ylab("Fitted Moth Estimates") +  
            theme_classic() +
            scale_y_continuous(breaks = seq(0, 220, 20), limits = c(0,220)) +
            scale_x_continuous(breaks = seq(0, 220, 20), limits = c(0,220)) +
            theme( panel.background = element_blank(),
                  plot.background = element_blank(),
                  panel.border = element_blank(),
                  legend.direction = "horizontal",
                  legend.position="bottom",
                  strip.text = element_text(size=16, face="bold"),
                  strip.background = element_blank(),
                  legend.key.size = unit(1,"line"),
                  legend.key.width = unit(3,"line"),
                  legend.text = element_text(size=12, face="bold"),
                  legend.title = element_text(size=16, face="bold"),
                  axis.title.x = element_text(face="bold", size = 24),
                  axis.title.y =element_text(face="bold", size = 24),
                  axis.text.x = element_text(face="bold", size=24, vjust=0.5, 
                                             hjust=0.5, angle=0),
                  axis.text.y = element_text(face="bold", size=24, vjust=0.5, 
                                             hjust=1, angle=0),
                  plot.title = element_text(size=22, face="bold"))

ggplot(plot.fitted, aes(Moths.cut, Fitted2)) +
         geom_jitter(shape = 1, size = 3, col = "gray70") +
         geom_abline(slope = 1, intercept = 0, alpha = 0.5, size = 1, col="red") +
         xlab("Observed Moth Counts") +
            ylab("Fitted Moth Estimates") +  
            theme_classic() +
            scale_y_continuous(breaks = seq(0, 220, 20), limits = c(0,220)) +
            scale_x_continuous(breaks = seq(0, 220, 20), limits = c(0,220)) +
            theme( panel.background = element_blank(),
                  plot.background = element_blank(),
                  panel.border = element_blank(),
                  legend.direction = "horizontal",
                  legend.position="bottom",
                  strip.text = element_text(size=16, face="bold"),
                  strip.background = element_blank(),
                  legend.key.size = unit(1,"line"),
                  legend.key.width = unit(3,"line"),
                  legend.text = element_text(size=12, face="bold"),
                  legend.title = element_text(size=16, face="bold"),
                  axis.title.x = element_text(face="bold", size = 24),
                  axis.title.y =element_text(face="bold", size = 24),
                  axis.text.x = element_text(face="bold", size=24, vjust=0.5, 
                                             hjust=0.5, angle=0),
                  axis.text.y = element_text(face="bold", size=24, vjust=0.5, 
                                             hjust=1, angle=0),
                  plot.title = element_text(size=22, face="bold"))
```


## Mean Spatial Field  
Plot average spatial effect across whole dataset.
```{r  fig.width=11, fig.height=8}
space.df = Model.0$summary.random$Region1[1:995,c(1,2,4,6)]   #pull estimates from model.  There are two copies, we take the first
names(space.df) = c("Region", "Mean", "Q2.5", "Q97.5")  #rename

hex_grid.plot = hex_grid #make copy for plotting

hex_grid.plot$Spatial.eff = with(space.df,
                             Mean[match(
                                hex_grid.plot$Region,
                                         Region)])     
range(hex_grid.plot$Spatial.eff)


#Format for ggplot
hex_grid.plot@data$id = rownames(hex_grid.plot@data)
hex_df = fortify(hex_grid.plot, region = "id")
hex_df = left_join(hex_df, hex_grid.plot@data, by = "id")

ggplot(hex_df, 
        aes(long,lat, group=group, fill = Spatial.eff)) + 
        scale_fill_gradient2("Random Spatial Field",
                             low = "blue", mid = "white", high ="red", 
                             midpoint = 0, space = "rgb", guide = "colourbar",
                             breaks = seq(-6, 4, 2),
                             limits = c(-6, 4)) +
       geom_polygon(col="black", size = 0.1) + 
        xlab(" ") +
        ylab(" ") +
        coord_equal() + 
        theme(panel.grid.minor = element_blank(),
              panel.grid.major = element_blank(),
              panel.background = element_blank(),
              plot.background = element_blank(),
              panel.border = element_blank(),
              #legend.direction = "horizontal",
              legend.position="bottom",
              strip.text = element_text(size=16, face="bold"),
              strip.background = element_blank(),
              legend.key.size = unit(1,"line"),
              legend.key.width = unit(3,"line"),
              legend.text = element_text(size=10, face="bold"),
              legend.title = element_text(size=14, face="bold"),
              axis.title.x = element_blank(),
              axis.title.y =element_blank(),
              axis.text.x = element_blank(),
              axis.text.y =element_blank(),
              axis.ticks.x = element_blank(),
              axis.ticks.y =element_blank(),
              plot.title = element_text(size=22, face="bold"))
```


## Time Trend  
Plot Temporal trend from RW1 term. 
```{r  fig.width=11, fig.height=8}
trend.df = Model.0$summary.random$Year1[,c(1,2,4,6)]   #pull estimates from model
names(trend.df) = c("Index", "Mean", "Q2.5", "Q97.5")  #rename

trend.df$Year = 1990:2021 #Years

myspan = 0.5 #smoothness for ggplot
ggplot(trend.df, aes(Year, Mean)) +
        geom_hline(yintercept = 0, 
                   linetype = "solid",
                   col = "darkgray",
                   size = 0.5) +
        geom_smooth(col = "black", 
                  method = "loess",
                  span = myspan,
                  se = FALSE,
                  lwd = 1) +
        geom_smooth(data = trend.df, aes(Year, Q2.5), 
                    col = "grey40", 
                    method = "loess",
                    span = myspan,
                    se = FALSE,
                    linetype= "dashed",
                    size=0.1) +
        geom_smooth(data = trend.df, aes(Year, Q97.5), 
                    col = "grey40", 
                    method = "loess",
                    span = myspan,
                    se = FALSE,
                    linetype= "dashed",
                    size=0.1) +
        xlab(" ") +
        ylab("Temporal Trend (log)") +  
        theme_classic() +
        theme(axis.text=element_text(size=16),
              panel.background = element_blank(),
              plot.background = element_blank(),
              panel.border = element_blank(),
              legend.position = "none",
              strip.background = element_blank(),
              strip.text = element_text(face="bold", size = 22),
              axis.title.y = element_text(face="bold", size = 24),
              axis.text.x = element_text(face="bold", size=18, vjust=0.5, 
                                         hjust=0.5, angle=0),
              axis.title.x = element_text(face="bold", size = 26))
```

#Air Temp trend

```{r  fig.width=11, fig.height=8}
trend.df = Model.0$summary.random$range.grp[,c(1,2,4,6)]   #pull estimates from model
names(trend.df) = c("Index", "Mean", "Q2.5", "Q97.5")  #rename

trend.df$Year = 10:30 #Years

myspan = 0.5 #smoothness for ggplot
ggplot(trend.df, aes(Index, Mean)) +
        geom_hline(yintercept = 0, 
                   linetype = "solid",
                   col = "darkgray",
                   size = 0.5) +
        geom_smooth(col = "black", 
                  method = "loess",
                  se = FALSE,
                  lwd = 1) +
        geom_smooth(data = trend.df, aes(Index, Q2.5), 
                    col = "grey40", 
                    method = "loess",
                    se = FALSE,
                    linetype= "dashed",
                    size=0.1) +
        geom_smooth(data = trend.df, aes(Index, Q97.5), 
                    col = "grey40", 
                    method = "loess",
                    se = FALSE,
                    linetype= "dashed",
                    size=0.1) +
        xlab(" ") +
        ylab("Temporal Trend (log)") +  
        theme_classic() +
        theme(axis.text=element_text(size=16),
              panel.background = element_blank(),
              plot.background = element_blank(),
              panel.border = element_blank(),
              legend.position = "none",
              strip.background = element_blank(),
              strip.text = element_text(face="bold", size = 22),
              axis.title.y = element_text(face="bold", size = 24),
              axis.text.x = element_text(face="bold", size=18, vjust=0.5, 
                                         hjust=0.5, angle=0),
              axis.title.x = element_text(face="bold", size = 26))
```

```{r}

p <- autoplot(Model.0)
plot(p)

p2 <- autoplot(Model.1)
plot(p2)

Model.1$summary.fitted.values$


library(inlabru)  
plot_dat <- predict(Model.1,Model.hex)



mean <- Model.2$summary.fitted.values$mean[idat2]
quant0.025 <- Model.2$summary.fitted.values$`0.025quant`[idat2]
quant0.975 <- Model.2$summary.fitted.values$`0.975quant`[idat2]
  
  
Model.hex2 <- Model.hex %>% cbind(mean,quant0.025,quant0.975) 


Model.1$summary.fitted.values[,c(1,3,5)] %>% as_tibble()






ggplot(Model.hex2, aes(x=Air_temp, y=mean)) +
        geom_smooth(col = "black", 
                  method = "loess",
                  se = FALSE,
                  lwd = 1) +
        geom_smooth(data = Model.hex2, aes(Air_temp, quant0.025), 
                    col = "grey40", 
                    method = "loess",
                    se = FALSE,
                    linetype= "dashed",
                    size=0.1) +
        geom_smooth(data = Model.hex2, aes(Air_temp, quant0.975), 
                    col = "grey40", 
                    method = "loess",
                    se = FALSE,
                    linetype= "dashed",
                    size=0.1) +
        geom_rug(sides="b") +
        xlab("Air Temperature (c)") +
        ylab("Predicted yearly month count per region") +  
        theme_classic() +
        coord_cartesian(ylim=c(0,300)) +
        theme(axis.text=element_text(size=16),
              panel.background = element_blank(),
              plot.background = element_blank(),
              panel.border = element_blank(),
              legend.position = "none",
              strip.background = element_blank(),
              strip.text = element_text(face="bold", size = 22),
              axis.title.y = element_text(face="bold", size = 24),
              axis.text.x = element_text(face="bold", size=18, vjust=0.5, 
                                         hjust=0.5, angle=0),
              axis.title.x = element_text(face="bold", size = 26))















ggplot(plot_dat,aes(y=Fitted2,x=Air_temp)) + 
  geom_smooth(method="loess") + 
  ggpubr::theme_pubr() + 
  #geom_point()
  geom_rug(sides="b") + 
  coord_cartesian(ylim=c(0,100))


ggplot(Model.hex,aes(y=Fitted2,x=Precip)) + geom_smooth(method="loess") + ggpubr::theme_pubr() + geom_rug(sides="b") + coord_cartesian(ylim=c(0,300))


```

