
```{r}
setwd("D:/Helicoverpa")
```
```{r}

#Main modeling package
suppressMessages(library(INLA))
#install.packages("INLA",repos=c(getOption("repos"),INLA="https://inla.r-inla-download.org/R/stable"), dep=TRUE)

#supporting packages
suppressMessages(library(reshape2))
suppressMessages(library(raster))
suppressMessages(library(corrplot))
suppressMessages(library(rasterVis))
suppressMessages(library(rgeos))
suppressMessages(library(rgl))
suppressMessages(library(maptools))
suppressMessages(library(mapproj))
suppressMessages(library(spdep))
suppressMessages(library(ggplot2))
suppressMessages(library(ggthemes))
suppressMessages(library(GISTools))
suppressMessages(library(lattice))
suppressMessages(library(gridExtra))
suppressMessages(library(spatialEco))
suppressMessages(library(dplyr))
suppressMessages(library(sp))
suppressMessages(library(pals))
```

# Observations
```{r}
rm(list=ls())

Heli.df = fread("data/processed/INLA_data_mesh_20220426.csv", 
                  header = TRUE, sep=",") %>% as_tibble() %>%
  select(trap,season,mean_count,se_count,species,type,year,Latitude,Longitude,Air_temp,Precip,num_yearly_traps,num_yearly_trap_obs) %>%
  pivot_wider(names_from = season,
              values_from = c(Air_temp, Precip)) %>%
  filter(type=="data") #filtering out all the mesh data just to look at the actual data

#Heli.df$mean_count[is.na(Heli.df$mean_count)] <- 0


dim(Heli.df)
str(Heli.df)

range(Heli.df$year)
unique(Heli.df$species)

range(Heli.df$Latitude)
range(Heli.df$Longitude)

range(Heli.df$num_yearly_traps,na.rm=TRUE) #sampling effort: number of traps per year
range(Heli.df$num_yearly_trap_obs,na.rm=TRUE) #sampling effort: number of trap obs per year


range(Heli.df$mean_count,na.rm=TRUE)
length(which(is.na(Heli.df$se_count)))
length(which(is.na(Heli.df))) #NAs come from the mesh data
```


## Observations to Spatial Points
```{r}
LL84 = "+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0"

Heli.pnts = SpatialPointsDataFrame(Heli.df[,c("Latitude","Longitude")], Heli.df)
proj4string(Heli.pnts) = LL84

plot(Heli.pnts, col="red", cex=0.1)
```


# Option 1 (Global Mesh)

## Spatial Extents

Load and simplify world boundaries
```{r}
LL84 = "+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0"

nProj = "+proj=merc +a=6378137 +b=6378137 +lat_ts=0 +lon_0=0
         +x_0=0 +y_0=0 +k=1 +units=km +nadgrids=@null +no_defs"

Countries = maps::map("world", 
                fill = TRUE, 
                plot = FALSE)


IDs = sapply(strsplit(Countries$names, ":"), function(x) x[1])

LL84 = "+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0"

Countries = map2SpatialPolygons(Countries, IDs = IDs,
                                proj4string = CRS(projection(LL84)))

Countries$Names = names(Countries)
Countries.copy = Countries #for comparison
Countries = subset(Countries, Names != "Antarctica" & Names != "Greenland") #Remove Antarctica and Greenlan

Countries = gBuffer(Countries, width = 0) #fixes some topological issues

#Add a dataframe  
pid = sapply(slot(Countries, "polygons"), 
             function(x) slot(x, "ID"))

p.df = data.frame( ID=1:length(Countries), 
                   row.names = pid)

Countries = SpatialPolygonsDataFrame(Countries, p.df)
map.copy = Countries

#Remove small islands
# breaks into separate polgons
Countries = explode(Countries, sp=TRUE)

#calculate area
Countries$Area = gArea(Countries, byid = TRUE) #ignore projection warnings

# Keep big areas
Countries = subset(Countries, Area > 500) 
Countries = gBuffer(Countries, width = 1) #smooting and enlarging

plot(Countries.copy, main = "Original") 
plot(Countries, main = "Simplified")
plot(Heli.pnts, add=TRUE, col="red", cex=0.1)

```




Create a raster/grid version for mapping results
```{r}
blank.r = raster(res = 1, crs = proj4string(map.copy))

extent(blank.r) = extent(map.copy)

World.r = rasterize(map.copy, 
                    blank.r,
                    field = 0, #zero value in polgons
                    background = NA) #value in oceans

#View raster
plot(World.r)

#Point version for later mapping
grid.pnts = rasterToPoints(World.r, sp = TRUE)

grid.pnts@data = grid.pnts@data %>%
                    mutate(pLong = grid.pnts@coords[,1],
                           pLat = grid.pnts@coords[,2],
                           Set = "grid") %>%
                    select(-layer)
```


## Construct Global Mesh
```{r}
MaxEdge = 2.0 #controls edge lengths in triangulations

bdry = inla.sp2segment(Countries) #formats for INLA package

mesh2D = inla.mesh.2d(boundary = bdry, 
                     cutoff = 1, 
                     max.edge = MaxEdge,
                     min.angle = 21) 

plot(mesh2D) #Flat version

MeshLocs = cbind(mesh2D$loc[,1], mesh2D$loc[,2]) #Pull coordinates from flat 2D mesh above (nodes/vertices)
head(MeshLocs) #long/lat

#Converting to 3D coordinates scaled to 1 Earth Radius
xyz = as.data.frame(                
          inla.mesh.map(MeshLocs, #coords from flat version
                        projection = "longlat", 
                        inverse = TRUE)) #invert from long/lat

true.radius.of.earth = 6371 #1 Earth Radius
radius.of.earth = 1 #scale radius to 1

Global.mesh = inla.mesh.create(loc = xyz,
                         cutoff = 100/true.radius.of.earth, #maximum length to allow
                         refine=list(max.edge = 3500/true.radius.of.earth, min.angle = 26))


Global.mesh$n #number of nodes. More nodes equal more precsion, but also more computer processing/memory

plot(Global.mesh, rgl = TRUE, main = " ") #this code produces a pop-up window where you can click, zoom, and drag mesh in 3D

plot(Global.mesh, #Fancy version above, showing mesh nodes over land
     rgl=TRUE, 
     col="lightblue", 
     draw.vertices=TRUE, 
     draw.edges=TRUE)
```

## Link Observations to Mesh 
Matching observation locations to Global.mesh.
```{r}
locs = cbind(Heli.pnts$Longitude, Heli.pnts$Latitude) #get moth coords

locs = inla.mesh.map(locs, 
                     projection = "longlat", 
                     inverse = TRUE) #invert to 3D coords to match with mesh


A.matrix = inla.spde.make.A(Global.mesh, loc=locs) #create sparse matrix matching moth locations to geographic position on mesh
```

## Set Spatial Prior
This model setting controls the strength of spatial effect (correlation decay)
```{r}
spde.1 = inla.spde2.pcmatern(Global.mesh,
                             prior.range=c(0.9, 0.01), #First value is a spatial range (1 = 1 Earth Radius), second value is a probability
                             prior.sigma=c(1, 0.01))  #similar to above, error
                           

#create a spatial index for each node
Field.1 = inla.spde.make.index("Field.1", #arbitrary name for the spatial field
                                spde.1$n.spde) #number of mesh nodes.  just an idex sthe model knows to estimate for each location

str(Field.1) # just index 1 to number of nodes (9459)
```




## Nearest Neighbor Distances  
```{r}
#Need to project data to make more accurate distance measurements

P.pp = spTransform(Heli.pnts, nProj) #Copy and project

head(P.pp@coords) #new coordinates

Heli.pnts$NN = spatstat.geom::nndist(P.pp@coords, k = 1) #spatstat.geom can conflict with other packages, k =  1 is "first" nearest neighbor    

range(Heli.pnts$NN) # in km

Heli.pnts$NN.round = round(Heli.pnts$NN/10, 1) #scaling and rounding

range(Heli.pnts$NN.round) #will plug this into an rw in Model.1.

```




## Prepare Model Input  
Example predictors and random effect indices.  
```{r}
#make copy of observations
Model.in = Heli.pnts@data


#Convert "Years" to an integer index 1 to total years
range(Model.in$year)

Model.in$Year.int = as.integer(as.factor(as.character(Model.in$year)))
range(Model.in$Year.int) #41 years

#simple way of acounting for number of samples
Model.in$yearly_traps = Model.in$num_yearly_traps 
Model.in$yearly_trap_obs = Model.in$num_yearly_trap_obs 

range(Model.in$yearly_traps,na.rm=TRUE) 
range(Model.in$yearly_trap_obs,na.rm=TRUE) 


#Individual variation traps
#length(unique(Model.in$trap))
dim(Model.in)[1]

Model.in$Trap.ind = as.integer(as.factor(Model.in$trap))
range(Model.in$Trap.ind)

```



## Organize Data
Because spatial index (Field.1) is a different length than moth dataframe, need to convert to a list object.  
```{r}
My.lst = list(c(Field.1,  #Spatial Index
                list(intercept.1 = 1)), #Intercept
                list(Temp_gs = Model.in[,"Air_temp_growing_season"], 
                     Temp_os = Model.in[,"Air_temp_off_season"], 
                     Precip_gs = Model.in[,"Precip_growing_season"], 
                     Precip_os = Model.in[,"Precip_off_season"], 
                     yearly_traps = Model.in[,"yearly_traps"],   #Number of samples
                     yearly_trap_obs = Model.in[,"yearly_trap_obs"],   #Number of samples
                     NN =  Model.in[,"NN.round"], #Nearest Neighbor distance
                     Year = Model.in[,"Year.int"],   #Index for year order
                     Trap = Model.in[,"Trap.ind"], #Variability by trap location 
                     Species = Model.in[,"species"])) #Species variable



Model.in$Count.round = round(Model.in$mean_count, 0) #Response variable, rounded to integer
range(Model.in$Count.round,na.rm = TRUE) 
hist(Model.in$Count.round) #poisson-like, but very large values

Stack.1 = inla.stack(data = list(Y = Model.in$Count.round), #Response variable  
                                  A = list(A.matrix, 1), #spatial matrix    
                            effects = My.lst, #all variables of interest
                                tag = "moth.1") #arbitary label/name
```

## Run Model

### Simple model first, spatial effect only
```{r}
#Model formula
Frm.0 = Y ~ -1 + intercept.1 + # the -1 species we use a custom intercept
                  f(Field.1,   #spatial index
                    model=spde.1) #spatial prior

#theta0 = Model.0$internal.summary.hyperpar$mean
theta0 = c(0.2406323, 2.1144357) #previous run hyperparameters


Model.0 = inla(Frm.0, 
               data = inla.stack.data(Stack.1, spde=spde.1), 
               family = "poisson", 
               verbose = TRUE,
               control.predictor = list(
                                      A = inla.stack.A(Stack.1 ), 
                                      compute = TRUE, 
                                      link = 1), 
               control.mode = list(restart = TRUE, theta = theta0), #mean estimate from first run, speeds up re-runs
               control.compute=list(dic = TRUE, waic = TRUE)) 
```

Results summary
```{r}
summary(Model.0) #intercept and spatial field 
```
    
    
    
    
    
map spatial field
```{r}
pLoc = cbind(grid.pnts$pLong, grid.pnts$pLat) #coords

pLoc = inla.mesh.map(pLoc, 
                     projection = "longlat", 
                     inverse = TRUE) # convert to 3D projection to match with mesh

Ap = inla.spde.make.A(Global.mesh, loc=pLoc) #match to mesh

grid.pnts$Spatial.Field = drop(Ap %*% Model.0$summary.random$Field.1$mean)  #mean values for spatial field, add to grid pnts

#map to grid
Model.0.rf = rasterize(grid.pnts, 
                      World.r,
                      field = "Spatial.Field", 
                      background = NA) 
plot(Model.0.rf)
```
    
    
    
    
    
    
    
    
    
# Expanded Models

##Linear trend of air temp and precip
Adding additional variables.
```{r}
#Model formula
Frm.1 = Y ~ -1 + intercept.1 + # the -1 species we use a custom intercept (fixed effect)
                f(Field.1,   #spatial index
                   model=spde.1) + #spatial prior (random effect)
                f(Year.int,         #random effect for year
                   model="ar1", #autocorrelation between years
                   constr=TRUE) + 
                f(NN.round,   #Nearest Neighbor
                   model="rw1", 
                   constr=TRUE,
                   scale.model=TRUE) +
                f(yearly_traps,   #Nearest Neighbor
                   model="rw1", 
                   constr=TRUE,
                   scale.model=TRUE) +
                f(yearly_trap_obs,   #Nearest Neighbor
                   model="rw1", 
                   constr=TRUE,
                   scale.model=TRUE) +
                f(Trap.ind,   # iid random effect for trap variation
        				   constr=TRUE, 
        				   model="iid") +
                f(species,   # iid random effect for spp differences (like a factor)
        				   constr=TRUE, #center on zero
        				   model="iid") +
                Air_temp_growing_season + 
                Air_temp_off_season + 
                Precip_growing_season +
                Precip_off_season 

#theta1 = Model.1$internal.summary.hyperpar$mean
#theta1 = c(-2.3362336, 1.6831341, 0.5495259, 1.0261536, -3.7205470, -1.2528197, 0.5737816) #previous run


#Execute/run model
Model.1 = inla(Frm.1, 
               data = inla.stack.data(Stack.1, spde=spde.1), 
               family = "poisson", 
               verbose = TRUE,
               control.predictor = list(
                                      A = inla.stack.A(Stack.1 ), 
                                      compute = TRUE, 
                                      link = 1), 
               control.mode = list(restart = TRUE),# theta = theta1), #mean estimate from first run, speeds up re-runs
               control.compute=list(dic = TRUE, waic = TRUE)) 
```

Some Results
```{r}
summary(Model.1) #summary

Model.1$summary.hyperpar[4,] # Rho = correlation.  Years are correlated at about 0.77
```



##random walks of air temp and precip

```{r}


Model.in$range.precip_off <- inla.group(Model.in$Precip_off_season  , n = 50, method = "quantile")
Model.in$range.precip_growing <- inla.group(Model.in$Precip_growing_season , n = 50, method = "quantile")
Model.in$range.air_growing <- inla.group(Model.in$Air_temp_growing_season , n = 50, method = "quantile")
Model.in$range.air_off <- inla.group(Model.in$Air_temp_off_season, n = 50, method = "quantile")





My.lst_rw = list(c(Field.1,  #Spatial Index
                list(intercept.1 = 1)), #Intercept
                list(Temp_gs = Model.in[,"range.air_growing"], 
                     Temp_os = Model.in[,"range.air_off"], 
                     Precip_gs = Model.in[,"range.precip_growing"], 
                     Precip_os = Model.in[,"range.precip_off"], 
                     yearly_traps = Model.in[,"yearly_traps"],   #Number of samples
                     yearly_trap_obs = Model.in[,"yearly_trap_obs"],   #Number of samples
                     NN =  Model.in[,"NN.round"], #Nearest Neighbor distance
                     Year = Model.in[,"Year.int"],   #Index for year order
                     Trap = Model.in[,"Trap.ind"], #Variability by trap location 
                     Species = Model.in[,"species"])) #Species variable


Model.in$Count.round = round(Model.in$mean_count, 0) #Response variable, rounded to integer
range(Model.in$Count.round) 
hist(Model.in$Count.round) #poisson-like, but very large values

Stack_rw = inla.stack(data = list(Y = Model.in$Count.round), #Response variable  
                                  A = list(A.matrix, 1), #spatial matrix    
                            effects = My.lst_rw, #all variables of interest
                                tag = "moth.1") #arbitary label/name







#Model formula
Frm.2 = Y ~ -1 + intercept.1 + # the -1 species we use a custom intercept (fixed effect)
                f(Field.1,   #spatial index
                   model=spde.1) + #spatial prior (random effect)
                f(Year.int,         #random effect for year
                   model="ar1", #autocorrelation between years
                   constr=TRUE) + 
                f(NN.round,   #Nearest Neighbor
                   model="rw1", 
                   constr=TRUE,
                   scale.model=TRUE) +
                f(yearly_traps,   #Nearest Neighbor
                   model="rw1", 
                   constr=TRUE,
                   scale.model=TRUE) +
                f(yearly_trap_obs,   #Nearest Neighbor
                   model="rw1", 
                   constr=TRUE,
                   scale.model=TRUE) +
                f(Trap.ind,   # iid random effect for trap variation
        				   constr=TRUE, 
        				   model="iid") +
                f(species,   # iid random effect for spp differences (like a factor)
        				   constr=TRUE, #center on zero
        				   model="iid") +
                f(range.air_growing,   #Nearest Neighbor
                   model="rw1", 
                   constr=TRUE,
                   scale.model=TRUE) +
                f(range.air_off,   #Nearest Neighbor
                   model="rw1", 
                   constr=TRUE,
                   scale.model=TRUE) +
                f(range.precip_growing,   #Nearest Neighbor
                   model="rw1", 
                   constr=TRUE,
                   scale.model=TRUE) +
                f(range.precip_off,   #Nearest Neighbor
                   model="rw1", 
                   constr=TRUE,
                   scale.model=TRUE) 

#theta1 = Model.1$internal.summary.hyperpar$mean
#theta1 = c(-2.3362336, 1.6831341, 0.5495259, 1.0261536, -3.7205470, -1.2528197, 0.5737816) #previous run


#Execute/run model
Model.2 = inla(Frm.2, 
               data = inla.stack.data(Stack_rw, spde=spde.1), 
               family = "poisson", 
               verbose = TRUE,
               control.predictor = list(
                                      A = inla.stack.A(Stack_rw), 
                                      compute = TRUE, 
                                      link = 1), 
               control.mode = list(restart = TRUE),# theta = theta1), #mean estimate from first run, speeds up re-runs
               control.compute=list(dic = TRUE, waic = TRUE)) 
```


```{r}
Model.0$dic$dic
Model.1$dic$dic
Model.2$dic$dic
```





##random walks of air temp and precip per species

```{r}


Model.in$range.precip_off <- inla.group(Model.in$Precip_off_season  , n = 50, method = "quantile")
Model.in$range.precip_growing <- inla.group(Model.in$Precip_growing_season , n = 50, method = "quantile")
Model.in$range.air_growing <- inla.group(Model.in$Air_temp_growing_season , n = 50, method = "quantile")
Model.in$range.air_off <- inla.group(Model.in$Air_temp_off_season, n = 50, method = "quantile")

Model.in$species_rep1 <- as.integer(as.factor(Model.in$species))
Model.in$species_rep2 <- as.integer(as.factor(Model.in$species))
Model.in$species_rep3 <- as.integer(as.factor(Model.in$species))
Model.in$species_rep4 <- as.integer(as.factor(Model.in$species))


#as.numeric(levels(as.factor(Model.in$species)))[Model.in$species]

#Model.in$MySppReplicates_1 = as.integer(levels(as.factor(Model.in$species)))
#Model.in$MySppReplicates_2 = as.integer(levels(as.factor(Model.in$species)))
#Model.in$MySppReplicates_3 = as.integer(levels(as.factor(Model.in$species)))
#Model.in$MySppReplicates_4 = as.integer(levels(as.factor(Model.in$species)))

#MySppReplicates = as.integer(levels(as.factor(MyDataframe$Species)))

My.lst_rw = list(c(Field.1,  #Spatial Index
                list(intercept.1 = 1)), #Intercept
                list(Temp_gs = Model.in[,"range.air_growing"], 
                     Temp_os = Model.in[,"range.air_off"], 
                     Precip_gs = Model.in[,"range.precip_growing"], 
                     Precip_os = Model.in[,"range.precip_off"], 
                     yearly_traps = Model.in[,"yearly_traps"],   #Number of samples
                     yearly_trap_obs = Model.in[,"yearly_trap_obs"],   #Number of samples
                     NN =  Model.in[,"NN.round"], #Nearest Neighbor distance
                     Year = Model.in[,"Year.int"],   #Index for year order
                     Trap = Model.in[,"Trap.ind"], #Variability by trap location 
                     Species = Model.in[,"species"],
                     species_rep1 = Model.in[,"species_rep1"],
                     species_rep2 = Model.in[,"species_rep2"],
                     species_rep3 = Model.in[,"species_rep3"],
                     species_rep4 = Model.in[,"species_rep4"])) #Species variable


Model.in$Count.round = round(Model.in$mean_count, 0) #Response variable, rounded to integer
range(Model.in$Count.round) 
hist(Model.in$Count.round) #poisson-like, but very large values

Stack_rw = inla.stack(data = list(Y = Model.in$Count.round), #Response variable  
                                  A = list(A.matrix, 1), #spatial matrix    
                            effects = My.lst_rw, #all variables of interest
                                tag = "moth.1") #arbitary label/name







#Model formula
Frm.3 = Y ~ -1 + intercept.1 + # the -1 species we use a custom intercept (fixed effect)
                f(Field.1,   #spatial index
                   model=spde.1) + #spatial prior (random effect)
                f(Year.int,         #random effect for year
                   model="ar1", #autocorrelation between years
                   constr=TRUE) + 
                f(NN.round,   #Nearest Neighbor
                   model="rw1", 
                   constr=TRUE,
                   scale.model=TRUE) +
                f(yearly_traps,   #Nearest Neighbor
                   model="rw1", 
                   constr=TRUE,
                   scale.model=TRUE) +
                f(yearly_trap_obs,   #Nearest Neighbor
                   model="rw1", 
                   constr=TRUE,
                   scale.model=TRUE) +
                f(Trap.ind,   # iid random effect for trap variation
        				   constr=TRUE, 
        				   model="iid") +
                f(species,   # iid random effect for spp differences (like a factor)
        				   constr=TRUE, #center on zero
        				   model="iid") +
                f(range.air_growing,   #Nearest Neighbor
                   model="rw1",
                   replicate = species_rep1,
                   constr=TRUE,
                   scale.model=TRUE) +
                f(range.air_off,   #Nearest Neighbor
                   model="rw1",
                   replicate = species_rep2 ,
                   constr=TRUE,
                   scale.model=TRUE) +
                f(range.precip_growing,   #Nearest Neighbor
                   model="rw1",
                   replicate = species_rep3,
                   constr=TRUE,
                   scale.model=TRUE) +
                f(range.precip_off,   #Nearest Neighbor
                   model="rw1",
                   replicate = species_rep4, 
                   constr=TRUE,
                   scale.model=TRUE) 

#theta1 = Model.1$internal.summary.hyperpar$mean
#theta1 = c(-2.3362336, 1.6831341, 0.5495259, 1.0261536, -3.7205470, -1.2528197, 0.5737816) #previous run


#Execute/run model
Model.3 = inla(Frm.3, 
               data = inla.stack.data(Stack_rw, spde=spde.1), 
               family = "poisson", 
               verbose = TRUE,
               control.predictor = list(
                                      A = inla.stack.A(Stack_rw), 
                                      compute = TRUE, 
                                      link = 1), 
               control.mode = list(restart = TRUE),# theta = theta1), #mean estimate from first run, speeds up re-runs
               control.compute=list(dic = TRUE, waic = TRUE)) 
```




```{r}
Model.0$dic$dic
Model.1$dic$dic
Model.2$dic$dic
Model.3$dic$dic


Model.0$waic$waic
Model.1$waic$waic
Model.2$waic$waic
Model.3$waic$waic

```







### View Spatial Random Field
Model error, latencies, model residuals, etc.
```{r fig.width=8, fig.height=8}
range(values(Model.0.rf), na.rm=T)
rng = seq(-18, 0, 22)


mCols = (coolwarm(1000))

cr = colorRampPalette(c(mCols), 
         bias = 1.7, alpha = TRUE)


levelplot(Model.0.rf),
          margin = FALSE,
          xlab = NULL, 
          ylab = NULL, 
          maxpixels = 1e5,
          col.regions = cr, alpha.regions = 1, at = rng,
          colorkey = list(labels=list(#at=c(0.02, 0.25, 0.50, 0.75, 0.98),  
                                 #labels=c("0.00", "0.25", "0.50", "0.75", "1.00"), 
                                 fontface='bold', cex=1.5),
                                 labels=list(cex=18),
                                 space = "bottom"), 
          par.strip.text = list(col = "transparent"), #fontface='bold', cex=1.5
          par.settings = list(axis.line = list(col = "transparent"),
                              strip.background = list(col = 'transparent'), 
                              strip.border = list(col = 'transparent')),
          scales = list(draw=FALSE)) + 
  latticeExtra::layer(sp.polygons(map.copy , fill = "transparent", alpha = 1, lwd=0.75, col = "grey80"))
```
    

Nearest Neighbors   

Curve looks Kind of crazy, but looks like it helps some...
```{r fig.width=8, fig.height=8, warning=FALSE}
mic.df = as.data.frame(Model.3$summary.random$NN[,1:6])
names(mic.df) = c("ID", "Mean", "sd", "Q025", "Q50", "Q975") #The "ID" column are the distances

ggplot(mic.df, aes(ID*10, Mean)) + #Previously scaled by 10
        geom_smooth(method = "loess",
                    se = FALSE, col = "black", 
                    linetype= "solid") +
        geom_smooth(data = mic.df, aes(ID*10, Q025), 
                    method = "loess",
                    se = FALSE, col = "grey", 
                    linetype= "dashed") +
        geom_smooth(data = mic.df, aes(ID*10, Q975), 
                    method = "loess",
                    se = FALSE, col = "grey", 
                    linetype= "dashed") +
        geom_hline(yintercept = 0, 
                   linetype = "dotted",
                   col = "red",
                   size = 1) + 
        geom_vline(xintercept = 0, 
                   linetype = "dotted",
                   col = "red",
                   size = 1) +
        xlim(0,25) +
        xlab("Distance (km)") +
        ylab("Fine-scale spatial structure (log)") +
        theme_classic() +
         theme(plot.title = element_text(hjust = 0.5),
                             axis.title.y = element_text(face="bold", size=18),
                             axis.title.x = element_text(face="bold", size=18),
                             title = element_text(face="bold", size=18, hjust=0.5),
                             strip.text.x = element_text(face="bold", size = 14, colour = "black"),
                             axis.text.y = element_text(face="bold", size=14),
                             axis.text.x = element_text(face="bold", size=14, vjust=0.5))

```


Trap level variation  
Red points exclude 0 from credible interval.  
```{r fig.width=8, fig.height=8}
my.data = as.data.frame(Model.3$summary.random$Trap)
names(my.data) = c("ID", "Mean", "sd", "Q025", "Q50", "Q975", "mode", "kld")

my.data$Important = ifelse(my.data$Q025 < 0 & my.data$Q975 < 0, 1,
                     ifelse(my.data$Q025 > 0 & my.data$Q975 > 0, 1, 0)) #Check credible intervals

Important.set = my.data %>% filter(Important == 1)
NotImportant.set = my.data %>% filter(Important == 0)

ggplot(NotImportant.set, aes(ID, Mean)) + 
        geom_point(size=2, pch=1, col = "gray75") +
        geom_linerange(aes(ymin=Q025, ymax=Q975), colour="gray75", size = 0.5) +
        geom_point(data=Important.set, aes(ID, Mean), 
                   size=2, pch=19, col = "red") +
        geom_linerange(data=Important.set, aes(ymin=Q025, ymax=Q975), colour="black") +
        geom_hline(yintercept = 0, 
                    linetype = "dotted",
                       colour = "red",
                       size = 0.75) +
            theme_classic() +
                   xlab("Individual Trap") +
                   ylab("Trap variation (log)") + 
                    theme(plot.title = element_text(hjust = 0.5),
                             axis.title.y = element_text(face="bold", size=18),
                             axis.title.x = element_text(face="bold", size=18),
                             title = element_text(face="bold", size=18, hjust=0.5),
                             strip.text.x = element_text(face="bold", size = 14, colour = "black"),
                             axis.text.y = element_text(face="bold", size=14),
                             axis.text.x = element_text(face="bold", size=14, vjust=0.5))

```


Species Effect
```{r fig.width=8, fig.height=8}
Model.3$summary.random$Species #species iid levels

my.data = as.data.frame(Model.1$summary.random$species)
names(my.data) = c("ID", "Mean", "sd", "Q025", "Q50", "Q975", "mode", "kld")

ggplot(my.data, aes(ID, y=Mean)) + 
                geom_point(size=3, pch=1, col = "red") +
                geom_linerange(aes(ymin=Q025, ymax=Q975), colour="gray50") +
                geom_hline(yintercept = 0, 
                            linetype = "dotted",
                               colour = "red",
                               size = 0.75) +
                    theme_classic() +
                           xlab("Species") +
                           ylab("Species Effect (log)") + 
                            theme(plot.title = element_text(hjust = 0.5),
                             axis.title.y = element_text(face="bold", size=18),
                             axis.title.x = element_text(face="bold", size=18),
                             title = element_text(face="bold", size=18, hjust=0.5),
                             strip.text.x = element_text(face="bold", size = 14, colour = "black"),
                             axis.text.y = element_text(face="bold", size=14),
                             axis.text.x = element_text(face="bold", size=14, vjust=0.5))
```


map spatial field for expanded model
```{r}
pLoc = cbind(grid.pnts$pLong, grid.pnts$pLat) #coords

pLoc = inla.mesh.map(pLoc, 
                     projection = "longlat", 
                     inverse = TRUE) # convert to 3D projection to match with mesh

Ap = inla.spde.make.A(Global.mesh, loc=pLoc) #match to mesh

grid.pnts$Spatial.Field = drop(Ap %*% Model.1$summary.random$Field.1$mean)  #mean values for spatial field, add to grid pnts

#map to grid
Model.3.rf = rasterize(grid.pnts, 
                      World.r,
                      field = "Spatial.Field", #
                      background = NA) 

```

### Compare Spatial Random Field
```{r fig.width=8, fig.height=10}
Field.stack = stack(Model.0.rf, Model.1.rf)
names(Field.stack) = c("Expanded")


range(values(Field.stack), na.rm=T)
rng = seq(-24, 52, 0.01)


mCols = (coolwarm(1000))

cr = colorRampPalette(c(mCols), 
         bias = 1.7, alpha = TRUE)


levelplot(Field.stack,
          layout = c(1),
          margin = FALSE,
          xlab = NULL, 
          ylab = NULL, 
          #main = "Occurrence Probabilty",
          names.attr= c("Spatial Only", "Expanded"),
          maxpixels = 1e5,
          col.regions = cr, alpha.regions = 1, at = rng,
          colorkey = list(labels=list(#at=c(0.02, 0.25, 0.50, 0.75, 0.98),  
                                 #labels=c("0.00", "0.25", "0.50", "0.75", "1.00"), 
                                 fontface='bold', cex=1.5),
                                 labels=list(cex=18),
                                 space = "bottom"), 
          par.strip.text = list(col = "transparent"), #fontface='bold', cex=1.5
          par.settings = list(axis.line = list(col = "transparent"),
                              strip.background = list(col = 'transparent'), 
                              strip.border = list(col = 'transparent')),
          scales = list(draw=FALSE)) + 
  latticeExtra::layer(sp.polygons(map.copy , fill = "transparent", alpha = 1, lwd=0.75, col = "grey80"))
```


