---
title: "INLA_mesh_with_data_modeling"
author: "Douglas Lawton"
date: "4/26/2022"
output: html_document
---

```{r}
setwd("/Users/ddlawton/Dropbox (ASU)/Research/Postdoc/Helicoverpa_climatic_variability")
```
```{r}

#Main package

suppressMessages(library(tidyverse))
suppressMessages(library(splitstackshape))
suppressMessages(library(anytime))
suppressMessages(library(lubridate))
suppressMessages(library(ggpubr))
suppressMessages(library(ggridges))
suppressMessages(library(patchwork))
```



# Load in all the csv files and combine them into one file

```{r}

files <- list.files(path="data/processed/INLA_mesh_with_data_04262022",full.names = TRUE,pattern = ".csv")


data <- read_csv(files) %>% bind_rows()

#GEE somehow messed up the empty columns in the csvs that didnt have moth count data. I am going to manually fix it here. no biggie.

data <- data %>%
  mutate(mdn_cnt = NA,men_cnt=NA,nm_yr__=NA,nm_yrl_=NA,se_cont=NA,ttl_cnt=NA)



files <- list.files(path="data/processed/INLA_mesh_with_data_04262022/manually_fix",full.names = TRUE,pattern = ".csv")


data2 <- read_csv(files) %>% bind_rows() %>% rbind(data)

```


#Data cleaning


```{r}

#There are obs before the remotely sensed data is available -- filtering those out
#Also, I had to import this data as a shapefile into GEE which shortened the names -- I am renaming them here

data3 <- data2 %>% filter(year >= 1981) %>% 
  rename(continent = contnnt,hemisphere = hemsphr,median_count = mdn_cnt, mean_count = men_cnt,num_yearly_trap_obs = nm_yr__,
         num_yearly_traps = nm_yrl_,se_count = se_cont,total_count = ttl_cnt, year_before = yer_bfr) %>%
  select(!c(year_before,time_nd,tm_strt))


#Alright, now I am going to extract the coordinates from the .geo column

data4 <- data3 %>%
  mutate(coords = str_extract(.geo,'(?<=\\[).+?(?=\\])')) %>%
  separate(coords,into=c("Latitude","Longitude"),sep=",")


length(unique(factor(data4$rowid)))
# So GEE exports all the RS data as an array in each row (eg [0.001,0.005,0.006,etc])
# I need to extract the data and put it into a long format to throw it into the models.


data5 <- data4 %>%
  filter(soil_moist_1 != "[]") %>%
  rowid_to_column(var="rowid") %>%
  select(!c(`system:index`,.geo))

data6 <- cSplit(setDT(data5)[, lapply(.SD, gsub, pattern = "[][]", 
                          replacement = "")], names(data5), sep=",", fixed = FALSE, "long")
head(data6)

data7 <- data6 %>% 
  tidyr::fill(rowid,.direction = "down") %>%
  as_tibble() %>%
  group_by(rowid) %>%
  fill(continent,hemisphere,median_count,mean_count,num_yearly_trap_obs,num_yearly_traps,se_count,species,
       trap,total_count,type,year,Latitude,Longitude,.direction="down")

#Alright, now I am going to get the numbers into recognizable units and selecting for the final columns

head(data7$soil_temp_2)
data8 <- data7 %>%
  mutate(Air_temp  = Air_temp - 273.15,
         Precip = Precip * 1000,
         dates = anytime::anytime(dates/1000),
         soil_temp_1 = soil_temp_1 - 273.15,
         soil_temp_2 = soil_temp_2 - 273.15,
         avg_soil_temp = ((soil_temp_1) + (soil_temp_2)) / 2,
         avg_soil_moist = ((soil_moist_1) + (soil_moist_2)) / 2) %>%
  select(1:11,16:24)

data9 <- data8 %>% drop_na(avg_soil_temp)

```

#Data formatting

Next I want to format the environmental data into either off or on season climatic variables.I have taken into consideration that the hemisphere seasons are flipped. The assumption is that anything on or below the equator has an 'off' season which is totally opposite to that of the northern hemisphere.

```{r}

#First make the on/off seasons based off the vernal/autumnal equinoxes 

data10 <- data9 %>%
  mutate(
    doy = yday(dates),
    season = case_when(
      hemisphere == "Northern" & doy>=79 & doy<=263 ~ "growing_season",
      hemisphere == "Northern" & doy<=79 | doy>=263 ~ "off_season",
      hemisphere == "Southern" & doy>=263 | doy<=79 ~ "growing_season",
      hemisphere == "Southern" & doy<=263 & doy>=79 ~ "off_season",
  ))

#Next I am taking the average of the climatic variables by the seasonal parition for each unique observation.


data11 <- data10 %>%
  group_by(rowid,season) %>%
  summarize(
    continent = first(continent),
    hemisphere = first(hemisphere),
    median_count = first(median_count),
    mean_count = first(mean_count),
    num_yearly_trap_obs = first(num_yearly_trap_obs),
    num_yearly_traps = first(num_yearly_traps),
    se_count = first(se_count),
    species = first(species),
    total_count = first(total_count),
    type = first(type),
    year = first(year),
    Latitude = first(Latitude),
    Longitude = first(Longitude),
    Air_temp = mean(Air_temp),
    Precip = mean(Precip),
    avg_soil_temp = mean(avg_soil_temp),
    avg_soil_moist = mean(avg_soil_moist )
  )



```


# Basic data exploration

Okay everything looks fine. Lets do some basic viz on the data to confirm it looks good.

```{r}
final_dat <- data11 %>%
  mutate(species = case_when(
    species == NA ~ type,
    TRUE ~ species
  ))

final_dat %>%
  group_by(hemisphere,continent,season) %>% #The NAs in continent is from the mesh data
  summarize(n=n())

final_dat %>%
  group_by(hemisphere,continent,type) %>%  #The NAs in continent is from the mesh data
  summarize(n=n())  

final_dat %>%
  group_by(hemisphere,continent,species) %>%  #The NAs in continent and species is from the mesh data
  summarize(n=n())  

summary(final_dat)

final_dat %>%
  ungroup() %>%
  select(Air_temp,Precip,avg_soil_temp,avg_soil_moist,type) %>%
  pivot_longer(cols=1:4) %>%
  ggplot(aes(x=value,y=type)) +
  geom_density_ridges() +
  facet_wrap(~name,scales="free")



final_dat %>%
  ungroup() %>%
  select(Air_temp,Precip,avg_soil_temp,avg_soil_moist,type,season) %>%
  pivot_longer(cols=1:4) %>%
  ggplot(aes(x=value,y=season)) +
  geom_density_ridges() +
  facet_wrap(~name+type,scales="free")


final_dat %>%
  ungroup() %>%
  select(median_count,mean_count,num_yearly_trap_obs,num_yearly_traps,se_count,total_count,species) %>%
  pivot_longer(cols=1:6) %>%
  ggplot(aes(x=value,y=species)) +
  geom_density_ridges() +
  facet_wrap(~name,scales="free")

```



# Relationship between trapping effort and counts

```{r}
moth_data <- final_dat %>% filter(type=="data") %>% filter(season == "growing_season") #pick a season just to not have double points

mean_total <- ggplot(moth_data,aes(x=total_count,y=mean_count)) + 
  geom_point(pch=21) +
  geom_smooth(size=1.25,
              method = "gam", 
              formula = y ~ s(x, bs = "tp", k = 10)) +
  theme_pubr() +
  ggtitle("Yearly mean count x total count")

mean_num_traps <- ggplot(moth_data,aes(x=num_yearly_traps ,y=mean_count)) + 
  geom_point(pch=21) +
  geom_smooth(size=1.25,
              method = "gam", 
              formula = y ~ s(x, bs = "tp", k = 10)) +
  theme_pubr() +
  ggtitle("Yearly mean count x # of traps")


mean_num_trap_obs <- ggplot(moth_data,aes(x=num_yearly_trap_obs,y=mean_count)) + 
  geom_point(pch=21) +
  geom_smooth(size=1.25,
              method = "gam", 
              formula = y ~ s(x, bs = "tp", k = 10)) +
  theme_pubr() +
  ggtitle("Yearly mean count x # of trap observations")

(mean_total + mean_num_traps) /( mean_num_trap_obs + plot_spacer())

```

# Relationship between count and environmental variables

```{r}

air_temp <- ggplot(moth_data,aes(x=Air_temp,y=mean_count,color=species)) + 
  #geom_point(pch=21) +
  geom_smooth(size=1.25,
              method = "gam", 
              formula = y ~ s(x, bs = "tp", k = 10)) +
  theme_pubr() +
  ggtitle("Moth count x Air temp")

precip <- ggplot(moth_data,aes(x=Precip,y=mean_count,color=species)) + 
  #geom_point(pch=21) +
  geom_smooth(size=1.25,
              method = "gam", 
              formula = y ~ s(x, bs = "tp", k = 10)) +
  theme_pubr() +
  ggtitle("Moth count x Precip")


soil_temp <- ggplot(moth_data,aes(x=avg_soil_temp,y=mean_count,color=species)) + 
  #geom_point(pch=21) +
  geom_smooth(size=1.25,
              method = "gam", 
              formula = y ~ s(x, bs = "tp", k = 10)) +
  theme_pubr()  +
  ggtitle("Moth count x Soil temp")

soil_moist <- ggplot(moth_data,aes(x=avg_soil_moist,y=mean_count,color=species)) + 
  #geom_point(pch=21) +
  geom_smooth(size=1.25,
              method = "gam", 
              formula = y ~ s(x, bs = "tp", k = 10)) +
  theme_pubr() +
  ggtitle("Moth count x Soil moist")

(air_temp + precip) /( soil_temp + soil_moist)
```



Alright, I could keep on going with the data exploration, but I think I should stop here and get to starting to model this data. I am going to save the final data (final_data) as a csv to the processed folder to clear the work space.

```{r}
write.csv(final_dat,file="data/processed/INLA_data_mesh_20220426.csv")
```



#####