---
title: "April 11 data reformatting"
author: "Douglas"
date: "4/11/2022"
output: html_document
---

#main purpose

I am reconstructing the data base to have relevant time spans for the remotely sensed data extraction revovling around the equinoxes. This will include the actual data at the NA mesh data. Hopefully.


```{r}
setwd("/Users/ddlawton/Dropbox (ASU)/Research/Postdoc/Helicoverpa_climatic_variability")
```

```{r}

#Packages
suppressMessages(library(INLA))
suppressMessages(library(reshape2))
suppressMessages(library(raster))
suppressMessages(library(corrplot))
suppressMessages(library(rasterVis))
suppressMessages(library(rgeos))
suppressMessages(library(rgl))
suppressMessages(library(maptools))
suppressMessages(library(mapproj))
suppressMessages(library(spdep))
suppressMessages(library(ggplot2))
suppressMessages(library(ggthemes))
suppressMessages(library(GISTools))
suppressMessages(library(lattice))
suppressMessages(library(gridExtra))
suppressMessages(library(spatialEco))
suppressMessages(library(tidyverse))
suppressMessages(library(sp))
suppressMessages(library(pals))
suppressMessages(library(maps))
suppressMessages(library(raster))
suppressMessages(library(rnaturalearth))
suppressMessages(library(janitor))
suppressMessages(library(lubridate))
```


# loading in the observations and combining them all together
```{r}
us_dat <- read_csv("data/raw/US_dat/Hzea_dat.csv")
Baker_dat <- read_csv("data/processed/AUS_data/Baker_dat.csv")
myron_dat <- read_csv("data/processed/AUS_data/myron_data.csv")
sa_dat <- read_csv("data/processed/SA_dat/combined_sa_dat.csv")


se <- function(x) sqrt(var(x)/length(x))

# combining all data together
 
b_cleaned <- Baker_dat %>% 
  drop_na(mean_count) %>%
  select(Week,before,mean_count,Species,dataset) %>%
  rename(Year = before, count=mean_count) %>%
  mutate(Latitude = -30.206009806160633, Longitude = 149.59603761497965, Trap="Baker_dat")

m_cleaned<- myron_dat %>%
  mutate(Week = isoweek(Date),Year = year(Date)) %>%
  pivot_longer(cols=c(H_arm,H_punc),names_to = "Species",values_to = "count") %>%
  drop_na(count) %>%
  select(dataset,Latitude,Longitude,Week,Year,Species,count,Trap)
  
AUS_dat <- rbind(b_cleaned,m_cleaned) %>% mutate(continent = "Australia")

US_dat2 <- us_dat %>%
  select(woy,year,CEW_sum,location,longitude,latitude)%>%
  rename(Week = woy, Year = year,count = CEW_sum, Trap = location,Latitude=latitude,Longitude=longitude) %>%
  mutate(Species = "H_zea", dataset="US")  %>% mutate(continent = "North_America")


sa_dat2 <- sa_dat %>%
  select(year,h_arm,h_zea,date,Latitude,Longitude,Trap_ID) %>%
  pivot_longer(cols=c(starts_with("h_")),names_to = "Species",values_to = "Count") %>%
  mutate(year = year(date), Week = isoweek(date), dataset="SA", continent = "South_America",
         Species = factor(Species),Trap = (Trap_ID)) %>%
  select(!date) %>%
  group_by(year,Week,Trap,Species) %>%
  summarise(Count = sum(Count,na.rm=TRUE), Longitude = first(Longitude), Latitude = first(Latitude), dataset = first(dataset), continent = first(continent)) %>%
  clean_names()
  



dat <- rbind(AUS_dat,US_dat2) %>% 
  clean_names() %>%
  rbind(sa_dat2) %>%
  mutate(
  species = case_when(
    species == "arm" ~ "H_arm",
    species == "punc" ~ "H_punc",
    species == "h_arm" ~ "H_arm",
    species == "h_punc" ~ "H_punc",
    species == "h_zea" ~ "H_zea",
    TRUE ~ species
  )
) 


unique(factor(dat$species))

unique(factor(dat$continent))


```


# Plotting to ensure all the points are on land

```{r}

world <- ne_countries() %>% st_as_sf()

ggplot(data=world) +
  geom_sf(aes(geometry=geometry)) +
  geom_point(data=dat,aes(x=longitude,y=latitude),pch=21,size=0.75) +
  theme_void()

#Everything looks fine. The south american dataset will be added shortly.


```


```{r}

# Summarizing to the year

yearly_dat <- dat %>%
  group_by(year,trap,species) %>%
  summarize(`mean count` = mean(count),median_count = median(count), total_count = sum(count),
            se_count = se(count),Latitude = first(latitude),Longitude = first(longitude), continent = first(continent),num_yearly_trap_obs = length(trap)) %>%
  group_by(year) %>%
  mutate(num_yearly_traps = length(unique(trap)),type = "data", Hemisphere = case_when(Latitude <= 0 ~ "Southern", Latitude > 0 ~ "Northern")) 


```


# constructing the global mesh

Load and simplify world boundaries
```{r}
LL84 = "+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0"

nProj = "+proj=merc +a=6378137 +b=6378137 +lat_ts=0 +lon_0=0
         +x_0=0 +y_0=0 +k=1 +units=km +nadgrids=@null +no_defs"

Countries = maps::map("world", 
                fill = TRUE, 
                plot = FALSE)


IDs = sapply(strsplit(Countries$names, ":"), function(x) x[1])

LL84 = "+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0"

Countries = map2SpatialPolygons(Countries, IDs = IDs,
                                proj4string = CRS(projection(LL84)))

Countries$Names = names(Countries)
Countries.copy = Countries #for comparison
Countries = subset(Countries, Names != "Antarctica" & Names != "Greenland") #Remove Antarctica and Greenlan

Countries = gBuffer(Countries, width = 0) #fixes some topological issues

#Add a dataframe  
pid = sapply(slot(Countries, "polygons"), 
             function(x) slot(x, "ID"))

p.df = data.frame( ID=1:length(Countries), 
                   row.names = pid)

Countries = SpatialPolygonsDataFrame(Countries, p.df)
map.copy = Countries

#Remove small islands
# breaks into separate polgons
Countries = explode(Countries, sp=TRUE)

#calculate area
Countries$Area = gArea(Countries, byid = TRUE) #ignore projection warnings

# Keep big areas
Countries = subset(Countries, Area > 500) 
Countries = gBuffer(Countries, width = 1) #smooting and enlarging

plot(Countries.copy, main = "Original") 
plot(Countries, main = "Simplified")
plot(Heli.pnts, add=TRUE, col="red", cex=0.1)
```

Create a raster/grid version for mapping results
```{r}
blank.r = raster(res = 1, crs = proj4string(map.copy))

extent(blank.r) = extent(map.copy)

World.r = raster::rasterize(map.copy, 
                    blank.r,
                    field = 0, #zero value in polgons
                    background = NA) #value in oceans

#View raster
plot(World.r)

#Point version for later mapping
grid.pnts = rasterToPoints(World.r, sp = TRUE)

grid.pnts@data = grid.pnts@data %>%
                    as_tibble() %>%
                    mutate(pLong = grid.pnts@coords[,1],
                           pLat = grid.pnts@coords[,2],
                           Set = "grid") %>%
                    dplyr::select(-layer)
```


## Construct Global Mesh and make a dataframe of points to get NA data

```{r}
MaxEdge = 2.0 #controls edge lengths in triangulations

bdry = inla.sp2segment(Countries) #formats for INLA package

mesh2D = inla.mesh.2d(boundary = bdry, 
                     cutoff = 1, 
                     max.edge = MaxEdge,
                     min.angle = 21) 

plot(mesh2D) #Flat version

MeshLocs = cbind(mesh2D$loc[,1], mesh2D$loc[,2]) %>% as_tibble() %>%
  rename(Longitude = V1, Latitude = V2) %>%
  mutate(type = "mesh", Hemisphere = case_when(Latitude <= 0 ~ "Southern", Latitude > 0 ~ "Northern")) %>%
  rowid_to_column(var="rowid") %>%
  mutate(ID = paste0(type,"_",rowid)) %>%
  select(!rowid)

years <- rep(seq(1981,2020,by=1),times=10471)


meshids <- rep(MeshLocs$ID,each=40) %>% cbind(years) %>%
  as_tibble() %>%
  rename(ID = ".") %>%
  left_join(MeshLocs,by="ID")


head(meshids)


```


# Now combining the Helicoverpa data frame with the empty mesh data

```{r}

meshids_bind <- meshids %>%
  rename(trap = ID,year=years) %>%
  mutate(continent = NA,num_yearly_trap_obs = NA,num_yearly_traps = NA, species = NA,
         `mean count` = NA,median_count = NA,se_count = NA, year=as.numeric(year),total_count = NA)

combined <- yearly_dat %>% rbind(meshids_bind)  %>% clean_names() %>%
  drop_na(year) %>%
  mutate(year_before = year - 1 ,
    time_start = case_when(
      hemisphere == "Southern" ~ paste0(year_before,"-03-20"),
      hemisphere == "Northern" ~ paste0(year_before,"-09-22")),
    time_end = case_when(
      hemisphere == "Southern" ~ paste0(year,"-03-20"),
      hemisphere == "Northern" ~ paste0(year,"-09-22"))) 

summary(combined)
```

# write out CSV for use in google earth engine

```{r}
write.csv(combined,file="data/processed/INLA_modeling_mesh_with_data_20220411.csv",row.names = FALSE)

# there is something weird happening with the import of the CSV. I am uploading it as a shapefile

Heli.pnts = SpatialPointsDataFrame(combined[,c("longitude","latitude")], combined) %>% st_as_sf()
plot(Heli.pnts)

dat_sf <- combined %>%
  st_as_sf(.,coords=c("longitude","latitude"),crs=4326)


view(dat_sf %>% filter(continent == "South_America"))

dat_sf2 <- dat_sf %>%
  mutate(before_year = year - 1,
        time_end = paste0(year,"-09-","30"),
         time_start  = paste0(before_year,"-02-","01")) %>%
  filter(before_year >= 1981) %>%
  select(!before_year) 
summary(dat_sf2)


st_write(dat_sf, "data/processed/ugly_shapefile/Heliocoverpa_data.shp")


```

